\label{chapter:chap5}
\begin{miniabstract}
Another crucial extension to the multi-agent policy would be the ability to infer the hidden state via its local information. We will consider the cooperative case in which the agent's action is available as part of \correctquote{public observation}. The main observation from \cite{nayyar2013decentralized, foerster2018bayesian} is that we can turn a partially observable multi-agent problem into single-agent POMDP by considering the public agent that forms a belief of the state given histories of public observation. Given its belief over the state, the public agent constructs a placeholder policy that acts according to each agents' individual observation. By having the same placeholder policy for all agents, we ensure that all the agents are coordinated correctly. We will start by introducing the Public Belief MDP framework. Then we introduce Variational Sequential Monte Carlo (VSMC) \cite{le2017auto, maddison2017filtering, naesseth2017variational}, which is a basis to Deep Variational Reinforcement Learning (DVRL) \cite{igl2018deep} - an algorithm that captures the belief overstate. We will follow \cite{shvechikovjoint} on the control as a probabilistic framework interpretation of DVRL to construct our multi-agent algorithms.
\end{miniabstract}

\section{Public Belief MDP}
\input{chapters/chapter05/contents/01_pub_belief_MDP}

\section{Variational Sequential Monte Carlo and Deep Variational Reinforcement Learning}
\input{chapters/chapter05/contents/02_VSMC_DVL}

\section{Probabilistic Public Agent}
\input{chapters/chapter05/contents/03_Prob_Pub_Belief}

\section{Implementation}
\input{chapters/chapter05/contents/04_Implementation}

\section{Conclusion}
In the last chapter, we consider the problem of representing a belief of others. We relaxed the problem by allowing only a cooperative setting, which leads us to public-belief MDP formulation. By having this formulation, we can turn the problem into a POMDP problem, which can be solved via VSMC and control as an inference framework. We proposed an algorithm that would allow us to perform a public belief update and using this belief to act.